---
title: "Generalized linear regression to Evaluate renal failure"
output:
  html_document:
    df_print: paged
---

```{r}
library(readxl)
path_data<-"BasedatosRegS.xlsx"
df <- read_excel(path_data,col_names=TRUE)
important_base_columns <-  c('year', 'month', 'Identificacion',
       '1_Edad', '2_Sexo', '47_Peso_en_Kilogramos', '48_Talla_en_metros', '49_IMC',
       '50_Presion_arterial_sistolica', '51_Presion_arterial_diastolica',
       'EstadioRenal1', '61_Creatinina2_mg_dl',
       '62_BUN2_mg_d', '63_Albúmina2',
       '64_Glucemia2_mg_dl', '65_HbA1C2', '66_Calcio_mg_dl',
       '69_Fosforo_mEq_L_l', '70_PTH_mg_dl', '71_Colesterol_Total_mg_dl',
       '71.2 LDL colest', '72_Colesterol_HDL_mg_dl', '74_Hemoglobina_mg_dl',
       'creatinina2_suero_exam',
       'corrected_calculated_tfg', 'calculated_EstadioRenal2')#'Proteinuria g/dL' missing the 0
colnames(df)

```
# Testing normality

```{r}
vec<-colnames(df)
importantvecs<-c("1.Proteinuria1.g/L","2.Microalbuminuria1","3.HbA1C.1","4.Creatinuria.1","5.Creatinina.1","6.Glucemia1","7.Edad", "8.Sexo","9.Nivel.ingresos","11.Estado.civil","13.Peso.Kg","14.Talla.m","15.IMC","16.Presion.Art.Sistolica","17.Presion.Art.Diastolica","19.Creatinina2.mg.dl","20.BUN1.mg.d","21.Albúmina2","22.Glucemia2.mg.dl","23.HbA1C2","24.Calcio.mg.dl","25.Fosforo.mEq.L.l","26.PTH.mg.dl","27.ColesterolTotal.mg.dl","28.LDLcolest","29.Colesterol.HDL.mg.dl","30.Hemoglobina.mg.dl","31.RCV.Rel.AlbuminuriaCreatinuria","36.creatinina3.mgdl","39.Proteinuria2","40.Proteinuria3","42.Microalbuminuria2","44.Microalbuminuria3","45.ParcialOrina","tfg_before","tfg_change","tfgD_per_month_before","46.Corrected.TFG")

vecsImp<-c("1.Proteinuria1.g/L","2.Microalbuminuria1","3.HbA1C.1","4.Creatinuria.1","5.Creatinina.1","6.Glucemia1","7.Edad", "8.Sexo","9.Nivel.ingresos","11.Estado.civil","13.Peso.Kg","14.Talla.m","15.IMC","16.Presion.Art.Sistolica","17.Presion.Art.Diastolica","19.Creatinina2.mg.dl","20.BUN1.mg.d","24.Calcio.mg.dl","25.Fosforo.mEq.L.l","26.PTH.mg.dl","27.ColesterolTotal.mg.dl","28.LDLcolest","29.Colesterol.HDL.mg.dl","30.Hemoglobina.mg.dl","31.RCV.Rel.AlbuminuriaCreatinuria","36.creatinina3.mgdl","19.Creatinina2.mg.dl","45.ParcialOrina","tfg_before","tfg_change","tfgD_per_month_before","46.Corrected.TFG")

importantvecsNum<-c("Identificacion",'1.Proteinuria1.g/L',"2.Microalbuminuria1","3.HbA1C.1","4.Creatinuria.1","5.Creatinina.1","6.Glucemia1","7.Edad","13.Peso.Kg","14.Talla.m","15.IMC","16.Presion.Art.Sistolica","17.Presion.Art.Diastolica","19.Creatinina2.mg.dl","20.BUN1.mg.d","21.Albúmina2","22.Glucemia2.mg.dl","23.HbA1C2","24.Calcio.mg.dl","25.Fosforo.mEq.L.l","26.PTH.mg.dl","27.ColesterolTotal.mg.dl","28.LDLcolest","29.Colesterol.HDL.mg.dl","30.Hemoglobina.mg.dl","31.RCV.Rel.AlbuminuriaCreatinuria","36.creatinina3.mgdl","tfg_before","tfg_change","tfgD_per_month_before","46.Corrected.TFG")# Only one sample in this columns "39.Proteinuria2","40.Proteinuria3","42.Microalbuminuria2","44.Microalbuminuria3"
cat<-setdiff(vecsImp,importantvecsNum)
ndf <- df[,importantvecsNum]


```
# only LDL colesterol is normal, therefore linear models shouldn't be implemented

```{r}
#sapply(ndf, class)# Every column is numeric
colSums(!is.na(ndf))

```


# GEneralized linear models
#seeing the response variable and estimateits family as first aproximation

```{r}
library(fitdistrplus)
 #x <- na.omit(df$"46.Corrected.TFG")
x <- df$"46.Corrected.TFG"[!is.na(df$"46.Corrected.TFG")]
plotdist(x, histo=TRUE, demp=TRUE)
descdist(x)
```
# It's like a gamma and a lognormal, but lets see better

# Fitting which distribution goes with the response variable
https://stats.stackexchange.com/questions/374452/family-of-glm-represents-the-distribution-of-the-response-variable-or-residuals

```{r}

library(gamlss.dist)
library(gamlss.add)

fit <- fitDist(x,try.gamlss = TRUE)

summary(fit)
fit$fits

```

# Preparing the data for a fit Imputing NA values with random forests:

```{r}
library(mice)

dfTFG<-ndf[complete.cases(df$"46.Corrected.TFG"),]# Dataframe that is only complete TFGs
# Imputing the data with midastouch

colnames(dfTFG)<-c("Identificacion","Proteinuria_gL","Microalbuminuria_1","HbA1C_1","Creatinuria_1","Creatinina_1","Glucemia_1","Edad","Peso_Kg","Talla_m","IMC","Presion_Art_Sistolica","Presion_Art_Diastolica","Creatinina2_mg_dl","BUN1_mg_d","Albumina_2","Glucemia2_mg_dl","HbA1C2","Calcio_mg_dl","Fosforo_mEq_L_l","PTH_mg_dl","ColesterolTotal_mg_dl","LDLcolest","Colesterol_HDL_mg_dl","Hemoglobina_mg_dl","RCV_Rel_AlbuminuriaCreatinuria","creatinina3_mgdl","tfg_before","tfg_change","tfgD_per_month_before","Corrected_TFG")
attach(dfTFG)
imputed_data <-  mice(dfTFG, method="rf",print=FALSE)# Random forest method
full_data <- complete(imputed_data) 

#RCV_Rel_AlbuminuriaCreatinuria and tfgD_per_month_before cannot be imputed

completecolumns <-c("Identificacion","Proteinuria_gL","Microalbuminuria_1","HbA1C_1","Creatinuria_1","Creatinina_1","Glucemia_1","Edad","Peso_Kg","Talla_m","IMC","Presion_Art_Sistolica","Presion_Art_Diastolica","Creatinina2_mg_dl","BUN1_mg_d","Albumina_2","Glucemia2_mg_dl","HbA1C2","Calcio_mg_dl","Fosforo_mEq_L_l","PTH_mg_dl","ColesterolTotal_mg_dl","LDLcolest","Colesterol_HDL_mg_dl","Hemoglobina_mg_dl","creatinina3_mgdl","tfg_before","tfg_change","Corrected_TFG")
full_data <- full_data[,completecolumns]
```

# Generalized linear model with known distribution (It has to be an exponential function the response variable) as fitted above

```{r}
library(gamlss.dist)
library(gamlss.add)
library(gamlss)


regglm <- gamlss(`Corrected_TFG` ~`Creatinina_1`+ Proteinuria_gL+Microalbuminuria_1+HbA1C_1+Creatinuria_1+Glucemia_1+Edad+Peso_Kg+Talla_m+IMC+Presion_Art_Sistolica+Presion_Art_Diastolica+Creatinina2_mg_dl+BUN1_mg_d+Albumina_2+Glucemia2_mg_dl+HbA1C2+Calcio_mg_dl+Fosforo_mEq_L_l+PTH_mg_dl+ColesterolTotal_mg_dl+LDLcolest+Colesterol_HDL_mg_dl+Hemoglobina_mg_dl+creatinina3_mgdl+tfg_before+tfg_change, data=full_data, family=BCPEo)
summary(regglm)
#compare the coeficcients, they are a little bit different
regglm$coefficients
plot(regglm)
```
# Adding a random effect per person: The family adjusted doesn't work, so we use nu.fix=TRUE

```{r}
regglmR <- gamlss(`Corrected_TFG` ~`Creatinina_1`+ Proteinuria_gL+Microalbuminuria_1+HbA1C_1+Creatinuria_1+Glucemia_1+Edad+Peso_Kg+Talla_m+IMC+Presion_Art_Sistolica+Presion_Art_Diastolica+Creatinina2_mg_dl+BUN1_mg_d+Albumina_2+Glucemia2_mg_dl+HbA1C2+Calcio_mg_dl+Fosforo_mEq_L_l+PTH_mg_dl+ColesterolTotal_mg_dl+LDLcolest+Colesterol_HDL_mg_dl+Hemoglobina_mg_dl+creatinina3_mgdl+tfg_before+tfg_change+random(factor(Identificacion)), data=full_data, family=BCPEo, nu.fix=TRUE)
#plot(regglmR)# see the residuals
#wp(regglmR) # http://www.gamlss.com/wp-content/uploads/2013/01/gamlss-manual.pdf page 127

#compare the coeficcients, they are a little bit different

## Seeing correlations of the variables
```
# The asumptions of the mixed effect didn't workout maybe by trying another family it could work or transforming the variables

```{r}
plot(regglmR)# see the residuals
wp(regglmR)
```



# Generalized linear model with random mixed effects per person
```{r}
regglmMM <- gamlss(`Corrected_TFG` ~`Creatinina_1` + random(factor(Identificacion)) , data=full_data )
# ,family=BCPEo doesnt work
#family=BCPEo , opt="optim", numIter=100
summary(regglmMM)
#compare the coeficcients, they are a little bit different
regglmMM$coefficients
getSmo(regglmMM)
```
# To take out terms of the model with a ChiSquared test
```{r}
library(MASS)
full_data<- subset(full_data, select = -c(Identificacion) )
mod1<-gamlss(Corrected_TFG~., data=full_data, family=BCPEo)
summary(mod1)
#dropterm(mod1, test="Chisq") #  http://www.gamlss.com/wp-content/uploads/2013/01/gamlss-manual.pdf Page 138
#addterm(mod1, scope=~.+(Proteinuria_gL+Edad+Creatinina_1+tfg_before+Microalbuminuria_1+BUN1_mg_d)^2, test="Chisq") #adding interaction terms

# Final model
mod2<-gamlss(`Corrected_TFG` ~`Creatinina_1`+Proteinuria_gL+Microalbuminuria_1+Edad+Peso_Kg+Presion_Art_Sistolica+Presion_Art_Diastolica+Creatinina2_mg_dl+BUN1_mg_d+Albumina_2+Glucemia2_mg_dl+Calcio_mg_dl+Fosforo_mEq_L_l+PTH_mg_dl+tfg_before+Microalbuminuria_1:Creatinina_1+BUN1_mg_d:tfg_before+Edad:tfg_before+Edad:BUN1_mg_d, data=full_data, family=BCPEo, control = gamlss.control(n.cyc = 400))
#Testing if one takes them out and add interactions in one command

plot(mod2)# see the residuals and check for assumptions
wp(mod2) # check ranges of the validity of the model

# To force converging one can try many more iterations by  adding: control = gamlss.control(n.cyc = 200)
```

# To compare models
```{r}
GAIC.table(mod1, mod2) 
# Its better the new model
```

# One can even make a better model by transforming its independant variables with:
5.1 Cubic splines, the cs() function . . . . . . . . . . . . . . . . . . . . . . . . . . . 96
5.2 Varying coefficient, the vc() function . . . . . . . . . . . . . . . . . . . . . . . . 97
5.3 Penalized splines, the ps() function . . . . . . . . . . . . . . . . . . . . . . . . . 102
5.4 The loess function lo() . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . . 106
5.5 Fractional polynomials, the fp() function

# Validation metrics

```{r}
path_data<-"BasedatosRegVal.xlsx"
df2 <- read_excel(path_data,col_names=TRUE)
importantvecsNum<-c("Identificacion",'1.Proteinuria1.g/L',"2.Microalbuminuria1","3.HbA1C.1","4.Creatinuria.1","5.Creatinina.1","6.Glucemia1","7.Edad","13.Peso.Kg","14.Talla.m","15.IMC","16.Presion.Art.Sistolica","17.Presion.Art.Diastolica","19.Creatinina2.mg.dl","20.BUN1.mg.d","21.Albúmina2","22.Glucemia2.mg.dl","23.HbA1C2","24.Calcio.mg.dl","25.Fosforo.mEq.L.l","26.PTH.mg.dl","27.ColesterolTotal.mg.dl","28.LDLcolest","29.Colesterol.HDL.mg.dl","30.Hemoglobina.mg.dl","31.RCV.Rel.AlbuminuriaCreatinuria","36.creatinina3.mgdl","tfg_before","tfg_change","tfgD_per_month_before","46.Corrected.TFG")
ndf2 <- df2[,importantvecsNum]

dfTFG2<-ndf2[complete.cases(ndf2$"46.Corrected.TFG"),]
colnames(dfTFG2)<-c("Identificacion","Proteinuria_gL","Microalbuminuria_1","HbA1C_1","Creatinuria_1","Creatinina_1","Glucemia_1","Edad","Peso_Kg","Talla_m","IMC","Presion_Art_Sistolica","Presion_Art_Diastolica","Creatinina2_mg_dl","BUN1_mg_d","Albumina_2","Glucemia2_mg_dl","HbA1C2","Calcio_mg_dl","Fosforo_mEq_L_l","PTH_mg_dl","ColesterolTotal_mg_dl","LDLcolest","Colesterol_HDL_mg_dl","Hemoglobina_mg_dl","RCV_Rel_AlbuminuriaCreatinuria","creatinina3_mgdl","tfg_before","tfg_change","tfgD_per_month_before","Corrected_TFG")
print(paste0("Número total de pacientes en validación: ",length(unique(dfTFG2$Identificacion))))
attach(dfTFG2)
numbers<-nrow(dfTFG2)
dfTFG2<-rbind(dfTFG,dfTFG2)
imputed_data <-  mice(dfTFG2, method="rf",print=FALSE)# Random forest method
full_data2 <- complete(imputed_data) 
completecolumns <-c("Identificacion","Proteinuria_gL","Microalbuminuria_1","HbA1C_1","Creatinuria_1","Creatinina_1","Glucemia_1","Edad","Peso_Kg","Talla_m","IMC","Presion_Art_Sistolica","Presion_Art_Diastolica","Creatinina2_mg_dl","BUN1_mg_d","Albumina_2","Glucemia2_mg_dl","HbA1C2","Calcio_mg_dl","Fosforo_mEq_L_l","PTH_mg_dl","ColesterolTotal_mg_dl","LDLcolest","Colesterol_HDL_mg_dl","Hemoglobina_mg_dl","creatinina3_mgdl","tfg_before","tfg_change","Corrected_TFG")
response=full_data2[,c("Corrected_TFG")]
full_data2 <- full_data2[,completecolumns]

full_data2 <- tail(full_data2,n=numbers)
full_data2$Identificacion<-NULL
predicted <- predict(mod2, newdata=full_data2, type = "response")

predicted<-tail(as.vector(predicted),n=numbers)
response<-tail(as.vector(response),n=numbers)


library(caret)
library(Metrics)
library(MLmetrics)
# R native functions
MAE(predicted, response)
mse(predicted, response)
 
# caret package functions 
rmse(predicted, response)
R2_Score(predicted, response)
smape(response, predicted)
```







